---
title: 'Tidy Tuesday 2021 Week 7: Inequality Data'
author: "Sean Jackson"
date: "2/12/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
## initial settings
knitr::opts_chunk$set(
  comment = NA,
  echo = TRUE,
  error = TRUE,
  cache = FALSE,
  message = FALSE,
  dpi = 96,
  warning = FALSE
)

## width to use when printing tables etc.
options(
  width = 250,
  scipen = 100,
  max.print = 5000,
  stringsAsFactors = FALSE
)
```

<style>
.table {
  width: auto;
}
ul, ol {
  padding-left: 18px;
}
pre, code, pre code {
  overflow: auto;
  white-space: pre;
  word-wrap: normal;
  background-color: #ffffff;
}
</style>

For this weeks submission I created several animations. The comparative time series data provided is perfect for this type of visualization. 

```{r libraries}
library(tidyverse)
library(tidytuesdayR)
library(viridis)
library(hrbrthemes)
library(gganimate)
```

Load and clean the data for the visualizations:

```{r load and clean data}
income_mean <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_mean.csv')

# reorder income quintiles
income_mean$income_quintile_f <- factor(income_mean$income_quintile,
                                             levels = c("Top 5%", "Highest", "Second", "Middle", "Fourth","Lowest"))
```

Create line plots faceted by race and income group:

```{r, fig.width=10,fig.height=15}
income_mean %>%
  filter(year > 1996 & 
           race == c('White Alone','Black Alone','Hispanic','Asian Alone', 'All Races') &
           dollar_type == 'Current Dollars') %>%
  ggplot(., aes(x=year,y=income_dollars, colour = race)) +
  geom_line() +
  scale_y_continuous() +
  facet_grid(income_quintile_f ~ race, scales = "free") +
  theme_ipsum() +
  theme(legend.position="none") +
  labs(x="Year", y="Total",
       title = "Mean Income 1996-2016",
       subtitle = 'Facet Grid Line Plot by Income Percentile and Race',
       caption="SeanPJ.com")
```

