---
title: 'Tidy Tuesday 2021 Week 6: Historically Black Colleges and Universities'
author: "Sean Jackson"
date: "2/5/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
## initial settings
knitr::opts_chunk$set(
  comment = NA,
  echo = TRUE,
  error = TRUE,
  cache = FALSE,
  message = FALSE,
  dpi = 96,
  warning = FALSE
)

## width to use when printing tables etc.
options(
  width = 250,
  scipen = 100,
  max.print = 5000,
  stringsAsFactors = FALSE
)
```

<style>
.table {
  width: auto;
}
ul, ol {
  padding-left: 18px;
}
pre, code, pre code {
  overflow: auto;
  white-space: pre;
  word-wrap: normal;
  background-color: #ffffff;
}
</style>

For this weeks submission I created several animations. The comparative time series data provided is perfect for this type of visualization. 

```{r libraries}
library(tidyverse)
library(readxl)
library(glue)
library(viridis)
library(hrbrthemes)
library(gganimate)
```

Load and clean the data for the visualizations:

```{r data loading and cleaning}
hbcu_all <- read_excel("tabn313.20.xls", sheet = 1)

#make the names easier to work with
hbcu_all <- hbcu_all %>%
  rename('Public2Year' = `2-year - Public` ,
         'Private2Year' = `2-year - Private`,
         'Public4Year' = `4-year - Public`,
         'Private4Year' = `4-year - Private`)

hbcu_colleges_perc <- hbcu_all %>%
  select(Year,Public2Year,Private2Year,Public4Year,Private4Year) %>%
  pivot_longer(cols = c(Public2Year,Private2Year,Public4Year,Private4Year), 
               names_to = "College Type", 
               values_to = "Total")

hbcu_colleges_perc_area <- hbcu_colleges_perc %>%
  group_by(Year,`College Type`) %>%
  summarise(n = sum(Total)) %>%
  mutate(Percentage = n / sum(n)) %>%
  select(-n) 
```

First animation will show the attendance to each university type over time using facet_wrap plots:

```{r animation 1}
q <- ggplot(hbcu_colleges_perc, aes(x = Year, y = Total, fill = `College Type`)) +
  facet_wrap(~ `College Type`) +
  geom_area(alpha=0.6 , size=.5, colour="black") +
  scale_fill_viridis(discrete = T) +
  theme_ipsum() +
  theme(legend.position="bottom") +
  labs(x="Year", y="Total",
       subtitle="Annual Attendance By University Type 1976 to 2015",
       caption="SeanPJ.com") +
  ggtitle("Historically Black Colleges and Universities") +
  transition_reveal(Year) 

animate(q,end_pause = 10)
```

Next animation shows the total attendance counts for all HBCUs in a stacked area chart:

```{r animation 2}
t <- ggplot(hbcu_colleges_perc, aes(x = Year, y = Total, fill = `College Type`)) +
  geom_area(alpha=0.6 , size=.5, colour="black") +
  scale_fill_viridis(discrete = T) +
  scale_y_continuous(breaks = c(0,50000,100000,150000,200000,250000,300000,350000) )+
  theme_ipsum() +
  theme(legend.position="bottom") +
  labs(x="Year", y="Total",
       subtitle="Annual Attendance By University Type 1976 to 2015",
       caption="SeanPJ.com") +
  ggtitle("Historically Black Colleges and Universities") +
  transition_reveal(Year) 

animate(t,end_pause = 10)
```

And the final animation shows the proprtional percentage share of total attendance for each college type over time.

```{r animation 3}
p <- ggplot(hbcu_colleges_perc_area, aes(x = Year, y = Percentage*100, fill = `College Type`)) +
  geom_area(alpha=0.6 , size=.5, colour="black") +
  scale_fill_viridis(discrete = T) +
  theme_ipsum() +
  theme(legend.position="bottom") +
  labs(x="Year", y="Percent",
       subtitle="Annual Attendance By University Type 1976 to 2015",
       caption="SeanPJ.com") +
  ggtitle("Historically Black Colleges and Universities") +
  transition_reveal(Year)
  
animate(p, end_pause = 10)
```
